Do syntactic and semantic similarity lead to interference effects?\\ Evidence from self-paced reading and event-related potentials using German

### 2024-11-22 paper/revision2/ResponseToReviewers.tex

\myexternaldocument{elsarticle-template-harv}

\begin{document}
\hfill \today
%\section{}
%\subsection{}

\noindent \textbf{Prof.\ Jarrold}\\
\noindent \textbf{Associate Editor}\\
\noindent \textbf{Journal of Memory and Language}
\vskip 1em

\noindent \textbf{Response letter for Submission JML-24-17, 2nd round of revisions}
\vskip 2em
\noindent Dear Prof. Jarrold, 
\vskip 1em
with my co-authors Shravan Vasishth and Himanshu Yadav, I would like to submit the second revision of our article titled ``Do syntactic and semantic similarity lead to interference effects? Evidence from self-paced reading and event-related potentials using German''.

We are grateful for the thoughtful and constructive comments and suggestions from yourself and the reviewers, and feel that the quality of the manuscript has been greatly improved by the revisions.
To address the concern that we overstate the implications of our findings, we have clarified at various places in the abstract and discussion of our results that our findings only apply to the present design. Although we have added clarifying text in this second round of revisions (e.g., regarding the implications, the prolonged reading times starting at the distractor and the use of syntactic and semantic cues), the manuscript length could be shortened by four pages. We have removed repetitions and revised unnecessary verbose text passages (e.g., in the description of the SPR results).
We have made our OSF repository public to ensure general accessibility and have updated the link in the manuscript accordingly.

@@

\section*{Editor's comments} 
\subsection*{Comment E.1}
\begin{quote}
``Dear Dr. Schoknecht,

Thank you for submitting your revised manuscript to the Journal of Memory and Language. This has now been reviewed by the same three experts who reviewed the previous submission, and you can find their comments below.

You will see from these that all three reviewers appreciate the thoroughness with which you responded to the issues raised on the initial submission. Reviewers 1 and 3 raise some remaining points but recommend acceptance, while Reviewer 2 is less persuaded that your revisions have adequately addressed their concerns. Given this set of reviews, I have come to a `revise' decision because I would like you to have one more go at addressing the remaining points made by all three reviewers.

I do not anticipate sending out the next version for review again, but rather anticipate being in a position to accept it at that point. However, it's important to note that that is not guaranteed, and one reason I have come to a `revise' rather than `accept subject to minor revisions' is that, like Reviewer 1, I was unable to access your data via your OSF link. \textbf{It is very important to us at JML that data are readily accessible, and that accessibility needs to be in place before any potential acceptance decision}.''
\end{quote}

\subsubsection*{Response to E.1}
We have made the OSF repository public to grant general accessibility. We paste our data availability statement below.

\begin{quote}
    \Paste{link}
\end{quote}

@@

I do have some suggestions. In my view these are only minor. They mainly concern just the way things are presented or discussed. I list my comments by page.

p. 16: osf data etc.

I wanted to check those, but \textbf{when I clicked the link, it said I don't have permissions and that I need to request access}.''
\end{quote}

\subsubsection*{Response to R1.1}
We have made the OSF repository public to grant general accessibility (see also our Response to E.1).

@@

\begin{quote}
    \Paste{confound}
\end{quote}

\subsection*{Comment R1.4}
\begin{quote}
``p. 51: ``For more discussion of the role of structural retrieval cues, see Franck and Wagers (2020) and Arnett and Wagers (2017)."

\textbf{It would be good to cite Kush et al. (2015) here} (after all, a significant portion of Franck and Wagers, 2020, builds on Kush et al. 2015 to explore ways to deal with structural cues).

Kush et al. (2015) Relation-sensitive retrieval: Evidence from bound variable pronouns. Journal of memory and language.''
\end{quote}
\subsubsection*{Response to R1.4}
We have added the suggested reference on page \pageref{cite_kush}:

@@

\begin{quote}
    \Paste{cite_kush}
\end{quote}

\subsection*{Comment R1.5}
\begin{quote}
``p. 55: ``$\Psi(d(x_i, y), 0 , \delta)$, where $\Psi(.|\delta)$"

I was a bit confused here, since the formula  ``$\Psi(d(x_i, y), 0 , \delta)$ " did not have any  ``$\Psi(.|\delta)$" part.''
\end{quote}

\subsubsection*{Response to R1.5}
We have changed the expression to make it clearer. On page \pageref{modeling}, we now say:

\begin{quote}
    \Paste{modeling}
\end{quote}

@@

\subsection*{Comment R2.4}
\begin{quote}
``3. Misalignment of SPR and ERP results: The results from the self-paced reading (SPR) and event-related potential (ERP) methods do not align, with the only observed effect—semantic interference—pointing in opposite directions: inhibitory in SPR but facilitatory in ERP. Specifically, the authors report a reversed effect in the P600 component, showing a reduced P600 in conditions of higher semantic interference. This contradicts typical predictions, where the P600 is usually greater under high interference. The authors interpret this as a facilitatory interference effect (i.e., easier processing with high semantic interference). However, if this were the case, \textbf{we would also expect faster reaction times (RTs) in high semantic interference conditions, while the opposite is observed in the SPR data}.''
\end{quote}

\subsubsection*{Response to R2.4}
The SPR results and the ERP results in the N400 window both show inhibitory interference, i.e., more processing difficulty under high vs.\ low semantic interference, manifesting in longer reading times and a more negative N400. We explain below in our response to R2.5 that the reduced P600 for high interference might be explained by facilitatory interference because under high semantic interference whatever noun was retrieved can be easily integrated.

In general, SPR results cannot show the same complex results pattern that the ERP results showed because they provide just one measure per region. Additionally, only at the critical (or spill-over) region facilitated integration might be observed and in line with the reviews, we do not interpret the reading times of the later regions because these were overshadowed by the reading times differences starting at the distractor.

@@

Revised version: ``By contrast, Bayes factors provided either no evidence for or even evidence against syntactic interference in both spatiotemporal windows (BF10 $<$ 1.2). Similarly, Bayes factors provided either no evidence for or even evidence against the interaction in both spatio-temporal windows (BF10 $<$ 1.2)."

\textbf{Did the authors change the analyses? Why what was reported as anecdotal evidence under narrow priors is now reported as no evidence or evidence against?}''
\end{quote}
\subsubsection*{Response to R2.7}
Yes, all analyses were changed in the first round of revision. Firstly, the priors were changed (directional priors in the original version and symmetrical priors in the revised version). Secondly, the method to compute Bayes factors was changed (from bridge sampling to Savage-Dickey, see also our Response to E.3). This resulted in small numerical changes of the results, e.g., BF10 = 2.5 in the original analysis decreased to BF10 = 1.2 in the revised analysis. Furthermore in the original version of the manuscript, we were inconsistent in our description of small BFs, i.e., BF10 = 1.2. Since the first revision and in line with \cite{lee2014bayesian}, we now consistently interpret such small BFs as providing no evidence either way.

\subsection*{Comment R2.8}
\begin{quote}
``2. As a side note, on p. 58, the authors conclude that ``the parser searches for a subject that is within the same clause but uses the animacy cue without reference to the clause in which a noun appears." Does this mechanism truly seem plausible to the authors? What would this imply in practice? \textbf{Does it suggest that semantic and syntactic information are entirely encapsulated from each other?} If we were to set aside the issues I raised earlier, this would be one of the paper's central conclusions, yet the authors fail to provide a credible mechanism to support it.''
\end{quote}
\subsubsection*{Response to R2.8}

To clarify the assumed independence of syntactic and semantic processing, we have revised section 7.3, starting on page \pageref{synsem}. We have also renamed the section to `The use of syntactic and semantic information' (the former title was `Relation to other sentence processing accounts').

@@

The authors state, ``The fillers were less syntactically complex than the experimental items but had at least one embedded clause and generally provided more variety." \textbf{In what way did they provide more variety? In which respects? Both fillers and test items involve retrieval; what is the role of the fillers in this context?}''
\end{quote}
\subsubsection*{Response to R2.9}
The fillers included different types of subordinate clauses and (more) modifiers at different positions within the sentence. By deviating from the rigid structure of the critical items, they provided more variety in regard to the employed sentence structures. To avoid creating an obvious juxtaposition of very complex sentences (critical items) and very simple sentences (fillers), the fillers included long-distance dependencies and at least one subordinate clause. We have revised the description of the fillers in the manuscript on page \pageref{fillers} in the following way:

\begin{quote}
    \Paste{fillers}
\end{quote}

\subsection*{Comment R2.10}
\begin{quote}
``4. There are numerous repetitions throughout the paper. A prime example is the paragraph on ``hyp testing using BF" (p. 45), which is repeated almost verbatim in the ``Discussion" section (p. 47). This redundancy occurs several times, making the paper excessively and unnecessarily long. \textbf{The authors should streamline the content to enhance clarity and conciseness.}''
\end{quote}

\subsubsection*{Response to R2.10}
We have removed the repetitions to improve clarity and conciseness (see also Response to E.4).

### 2024-11-22 paper/revision2/elsarticle-template-harv.tex
Add files via upload

Reading times across the whole sentence are shown in Figure \ref{fig:whole_sentence}, separately for high \textcolor{blue}{(panels A and C)} and low \textcolor{blue}{(panels B and D)} syntactic interference because these conditions had partially different sentence structures.

>>>>

Reading times across the whole sentence are shown in Figure \ref{fig:whole_sentence}, separately for high \textcolor{blue}{(panel A)} and low \textcolor{blue}{(panel B)} syntactic interference because these conditions had partially different sentence structures.
